---
title: "Homework 1"
author: "Lisa Eisler"
date: "9/17/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1

In which we demonstrate the creation and manipulation of variable types and introduce coercion.

Here is a **code chunk** that creates hw1_df1 sampling from a _normal distribution_:
```{r}
hw1_df1 = tibble(
  normal = rnorm(8),
  normal_pos = normal > 0,
  name_char = c("My", "name", "is", "Lisa", "Danielle", "Eisler", "MD", "MS"),
  letter_factor = factor(c("Yes", "No", "Maybe", "Maybe", "No", "Yes", "Yes", "No"))
)
```

The mean of the numeric variable is `r mean(pull(hw1_df1, normal))`.
The mean of the logical variable is `r mean(pull(hw1_df1, normal_pos))`.
The mean of the character variable is `r mean(pull(hw1_df1, name_char))`.
The mean of the factor variable is `r mean(pull(hw1_df1, letter_factor))`.
I am unable to take the mean of character and factor variables.


Here is a **code chunk** that creates hw1_numeric_df1 sampling from a _normal distribution_, converting our original logical, character, and factor variables to numeric.

```{r, eval=FALSE}
 hw1numeric_df1 = tibble(
  normal = rnorm(8),
  normal_pos = as.numeric(normal > 0),
  name_char = as.numeric(c("My", "name", "is", "Lisa", "Danielle", "Eisler", "MD", "MS")),
  letter_factor = as.numeric(factor(c("Yes", "No", "Maybe", "Maybe", "No", "Yes", "Yes", "No")))
)
```

Logical variable values are convereted to numbers: True is converted to 1 and False is converted to 0. Factor variable values are converted to a number corresponding to level: A:1, B:2, C:3, etc. Character variable values are converted to NA.

This may explain why we are able to calculate the mean of a logical variable, even prior to converting it to numeric.

Lastly, here is a **code chunk** which creates hw1convert_df1, sampling from a _normal distribution_ and demonstrates conversion and nested conversion of logical variables, with column cross products.
```{r}
hw1convert_df1 = tibble(
  normal = rnorm(8),
  normal_pos_numeric = as.numeric(normal > 0),
  normal_pos_factor = as.factor(normal > 0),
  normal_pos_numeric_factor = as.numeric(as.factor(normal > 0)),
  product1 = normal*normal_pos_numeric,
  product2 = normal*normal_pos_factor,
  product3 = normal*normal_pos_numeric_factor
)

```

#Problem 2

Here is a **code chunk** that creates hw1_df2 sampling from a _normal distribution_:
```{r}
hw1_df2 = tibble(
  x = rnorm(500),
  y = rnorm(500),
  sum_greater1 = x + y > 1,
  sum_greater1numeric = as.numeric(x + y > 1),
  sum_greater1factor = as.factor(sum_greater1numeric)
)
```

The number of rows is
`r nrow(hw1_df2)`.
The number of columns is
`r ncol(hw1_df2)`.
The mean of x is 
`r mean(pull(hw1_df2, x))`.
The median of x is
`r median(pull(hw1_df2, x))`.
The standard deviation of x is
`r sd(pull(hw1_df2, x))`.
The proportion of cases for which x + y > 1 is to be determined.

Here is a scatter plot adding color using logical variable.
```{r yx_scatter1}
ggplot(hw1_df2, aes(x = x, y = y, color = sum_greater1)) + geom_point()
```

Here is a scatter plot adding color using numeric variable.
```{r yx_scatter2}
ggplot(hw1_df2, aes(x = x, y = y, color = sum_greater1numeric)) + geom_point()
```

Here is a scatter plot adding color using factor variable.
```{r yx_scatter3}
ggplot(hw1_df2, aes(x = x, y = y, color = sum_greater1factor)) + geom_point()
```